---
title: "Field_nutrients"
output: html_document
date: "2024-12-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Nutrients

```{r}
# Plots
MyTheme<-theme_bw() +  
theme(legend.position="top",
          plot.background=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5),
          panel.grid.major.y = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          legend.box.background = element_rect(),
          legend.title = element_blank(),
          panel.background =element_rect(fill = NA, 
                                         color = "black"))#+
  #guides(fill=guide_legend(nrow=2,byrow=TRUE), shape=guide_legend(nrow=3,byrow=TRUE))

Season_fill<-scale_fill_manual(values =
                           c("#2b83ba", "#abdda4",
                             "#990000", "#ff9900"))

Season_colour<-scale_colour_manual(values =
                          c("#2b83ba", "#abdda4",
                             "#990000", "#ff9900"))

Precipitation_fill<-scale_fill_manual(values =
                           c("#cc6633", "#66ccff"))

Precipitation_colour<-scale_colour_manual(values =
                           c( "#cc6633", "#66ccff"))

Zone_shapes4<- scale_shape_manual(values=c(21,23,24, 25))
Zone_shapes3<- scale_shape_manual(values=c(21,23,24))

```

# 1. Import and format carbon chemistry data

```{r, include=FALSE}
# 1. Get data
  FLK.data<-read.csv("WS_Data_998-2022.csv", header = T)
  summary(FLK.data)
  #head(FLK.data)

# 2. Format variables
  # Date/times
    FLK.data$Date<-as.Date(FLK.data$Date, format = "%m/%d/%y")
    #summary(FLK.data$Year)

    # FLK.data$Month<-factor(FLK.data$Month, 
    #                       levels = c("Apr", "May","Jun",
    #                                 "Jul", "Aug", "Sep",
    #                                 "Oct","Nov","Dec",
    #                                 "Jan", "Feb", "Mar"))
    # 
    # FLK.data$Month2<-factor(FLK.data$Month, 
    #                       levels = c("Jan", "Feb", "Mar",
    #                                  "Apr", "May","Jun",
    #                                 "Jul", "Aug", "Sep",
    #                                 "Oct","Nov","Dec"
    #                                 ))
    
  FLK.data$Region<-factor(FLK.data$Region, levels = c(
                          "BISCAYNE BAY", "UPPER KEYS", "MIDDLE KEYS", "LOWER KEYS"))
  
  FLK.data$Zone<-factor(FLK.data$Position, levels = c(
                        "INSHORE", "MID CHANNEL", 
                        "OFFSHORE", "OCEANIC"))
  
  FLK.data$Season<-factor(FLK.data$Season, levels = c(
                        "Winter", "Spring", "Summer", "Fall"))
#summary(FLK.data$Extreme)

# ## Seasonality (NOPE)
# -   Winter: Dec 20-March 19
# -   Spring: March 20-June 18
# -   Summer: June 19-Sep 20
# -   Fall: Sep 21-Dec 19
```

## Seasonality

-   Winter: Jan-Mar
-   Spring: Apr-Jun
-   Summer: Jul-Sep
-   Fall: Oct-Dec
-   Wet: May - October
-   Dry: November - April

# 2. Check data available

## 2.1 Locations (all samples)

```{r MapINformation, include=FALSE}
Locations <- FLK.data %>% select(Latitude,
                                Longitude, 
                                Region, 
                                Station)
Locations<-unique(Locations)

S.GPS.Sites <- plyr::ddply (FLK.data, .(Region, Zone, Station),summarise,
                Lat = mean (Latitude, na.rm = T), 
                Lon = mean (Longitude, na.rm = T),
                number = n())
#write.csv(WS.GPS.Sites, "FLK_results/1_meanSatationGPS.csv")

```

## 2.3 Number of samples

### By zone and location

```{r, echo=FALSE}
library(kableExtra)
Cases_Loc_Zone<-FLK.data %>% count(Region,
                                  Zone,
                                  sort = F)

Cases_Loc_Zone<-as.data.frame(Cases_Loc_Zone %>%
                   pivot_wider(names_from = Zone,
                   values_from = n))
#write.csv(Cases_Loc_Zone, "FLK_results/0_SampleS_Region_zone.csv")

kable(as.data.frame(Cases_Loc_Zone, format = "html"), 
      longtable = TRUE) %>% 
kable_styling(bootstrap_options = c("striped", "condensed", full_width = F), font_size = 11)
```

```{r}
NH4_plot<- ggplot(FLK.data) + MyTheme+
  geom_point(aes (Date, NH4, fill=Region, shape=Zone),
             alpha=0.6, size=.8)+
  # geom_smooth(aes(TDate, NH4, colour=Region),
  #             linetype=1, method = "lm", se=F)+
    geom_smooth(aes(Date, NH4),
               linetype=1, method = "lm", colour="black")+
  Zone_shapes4+
  Season_fill+
  guides(shape="none")+
  scale_y_continuous(limits = c(0, 10),
                     expand = c(0.01, 0.01),
                     name=expression(paste("NH4")))
NH4_plot + facet_grid(Region~Zone)
```


```{r}
PO4_plot<- ggplot(FLK.data) + MyTheme+
  geom_point(aes (Date, PO4, fill=Region, shape=Zone),
             alpha=0.6, size=.8)+
  # geom_smooth(aes(Date, PO4, colour=Region),
  #             linetype=1, method = "lm", se=F)+
    geom_smooth(aes(Date, PO4),
               linetype=1, method = "lm", colour="black")+
  Zone_shapes4+
  Season_fill+
  guides(shape="none")+
  scale_y_continuous(limits = c(0, 2),
                     expand = c(0.01, 0.01),
                     name=expression(paste("PO4")))
PO4_plot + facet_grid(Region~Zone)
```


```{r}
NO3_plot<- ggplot(FLK.data) + MyTheme+
  geom_point(aes (Date, NO3, fill=Region, shape=Zone),
             alpha=0.6, size=.8)+
  # geom_smooth(aes(Date, NO3, colour=Region),
  #             linetype=1, method = "lm", se=F)+
    geom_smooth(aes(Date, NO3),
               linetype=1, method = "lm", colour="black")+
  Zone_shapes4+
  Season_fill+
  guides(shape="none")+
  scale_y_continuous(limits = c(0, 10),
                     expand = c(0.01, 0.01),
                     name=expression(paste("NO3")))
NO3_plot + facet_grid(Region~Zone)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
